Performance comparison of Dask and Apache Spark on HPC systems \HL{for Neuroimaging}

### 2022-08-25 latex/paper.tex
[FIX] first draft to address second reviewers comments.

\documentclass[AMA,STIX1COL]{WileyNJD-v2}
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage[binary-units=true]{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{caption}
\usepackage{subcaption}

@@

\title{Performance comparison of Dask and Apache Spark on HPC systems}

@@

\abstract[Summary]{
	The general increase in data size and data sharing motivates the adoption of Big Data strategies in several scientific disciplines.
	However, while several options are available, no particular guidelines exist for selecting a Big Data engine.
	In this paper, we compare the runtime performance of two popular Big Data engines with
	Python APIs, Apache Spark, and Dask, in processing neuroimaging pipelines.
	Our experiments use three synthetic data-intensive applications to process
	the \SI{606}{\gibi\byte} BigBrain image and an actual pipeline to process data
	from thousands of anatomical images.
	We benchmark these applications on a dedicated HPC cluster running the Lustre file system while using varying
	combinations of the number of nodes, file size, and task duration.
	Our results show that although there are slight differences between Dask and Spark,
	the performance of the engines is comparable for data-intensive applications.
	However, Spark requires more memory than Dask, which can lead to 
	slower runtime depending on configuration and infrastructure.
	In general, the limiting factor was the data transfer time.
	While both engines are suitable for neuroimaging, more efforts need to be put
	to reduce the data transfer time and the memory footprint of applications.
}

@@

\section{Introduction}
The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO2018400, van2014human, ConpPortal}.
While standard neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipelines,
they lack Big Data strategies (i.e., in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
A previous study\cite{8752675} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
This work studies how performance benchmarks generalize across Big Data engines in an HPC environment.
In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC pipelines.

@@

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
Lustre is a reference file system for HPC environments, used in over 60 TOP100 supercomputers~{\cite{OpenSFS-lustre}}, offering parallel file accesses.

In a previous study\cite{8943502} we found no significant performance difference between Apache Spark and Dask for data-intensive applications.
However, using the network file system (NFS) led to an I/O bottleneck limiting application performance.
Another study~\cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
This same study found that the startup overhead time for Spark was more prominent than for Dask.
On the contrary, the work in~\cite{Mehta:17} claims that the startup overhead is more significant for Dask than for Spark.
This contradiction suggests that the startup overhead might vary depending on cluster configuration.
The latter study processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
On the contrary, the work in~\cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
However, that study used the Dask Bag API, which has performance limitations compared to other Dask APIs.

@@

The following section introduces the Lustre file system and the Big Data engines.
Next, we present the design of our benchmarking experiments.
We consider two types of data-intensive neuroimaging pipelines: high-resolution imaging and functional MRI studies for a cohort of subjects.
The pipelines we chose involve different compute and I/O patterns, e.g., map-reduce or map-only, requiring partial data in memory or the whole dataset.
Further sections present our results, discussion, and conclusion.

@@

Our system uses similar hardware as Compute Canada's B\'eluga cluster --- a representative cluster available to researchers in Canada --- and replicates as
close as possible its configuration.
This lets us compare Spark and Dask in the condition of a realistic
HPC system.

Both Spark and Dask are configured to have eight worker processes, each with eight threads.
Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
For Spark, the JVM heap space uses 10\% of workers' memory, and the location of the log is set on an NFS-mounted directory.
The worker workspace for both engines is located on Lustre due to the limited fast storage space on the worker nodes.
Other Dask configurations are left to default.
The Spark network timeout is set to \SI{600}{\second} and the executor heartbeat to \SI{120}{\second}.
A new Dask or Spark cluster was started and torn down for each experiment.
Dask \textit{2021.7.0} and Spark \textit{3.1.2} were used.

@@

\subsection{Applications}
This section describes the applications used to benchmark Dask and Spark.
We use three simple synthetic applications to have a deep understanding of the 
underlying behavior of the applications with simple I/O patterns, namely
\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
Also, two neuroimaging applications are used to study more realistic 
applications, namely \textit{Kmeans} and \textit{BIDS App Example}{\cite{gorgolewski2017bids}}.

To have a fair comparison between the engines, the code base for the core functions of the applications is the same for both engines.
This allows us to reduce the performance difference from different implementations.

@@

\subsubsection{Histogram}
As our third application, we calculate the histogram of the BigBrain image. 
The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre (Fig.~\ref{fig:graph-histogram}).
This map-reduce application has a very high read over write ratio.
Moreover, this application requires shuffling, albeit of a limited amount of data. 
The amount of inter-worker communication is in-between the increment and multi-increment applications.

\begin{figure}[h]
	\centering
	\begin{minipage}{.45\textwidth}
		\centering
		\includegraphics[width=0.45\linewidth]{figures/increment.png}
		\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
		\label{fig:graph-increment}		
	\end{minipage}
	\begin{minipage}{.45\textwidth}
		\centering
		\includegraphics[width=0.45\linewidth]{figures/histogram.png}
		\caption{Task graph for Histogram with 3 BigBrain blocks.}
		\label{fig:graph-histogram}			
	\end{minipage}
\end{figure}

\subsubsection{Kmeans}
For our fourth application, we apply Kmeans clustering to the voxel
intensities of the BigBrain image. We set the number of clusters to 3. The
application starts by reading the image blocks, combining all voxels in a
1-D array, and choosing initial centroids using the min, max, and
intermediate values. It assigns each voxel to its closest centroid and
updates each centroid with the average of its assigned voxels. It repeats
the assignment and updates steps for a configurable number of iterations.
Finally, the voxels of the image blocks are classified and written back to
the file system. Updating the centroids involves substantial data
communication between the workers.
The computation to find the closest centroids for each voxels scales with the number of centroids.
It can be computed in parallel at the cost of higher memory or sequentially to reduce memory usage.
We decided to compute the nearest centroids in parallel since we use only three centroids.
Due to resources limitation on our infrastructure, the Kmeans experiments
are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).

\subsubsection{BIDS App Example}
Our fifth application is the BIDS App Example, a neuroimaging pipeline to
measure brain volume from MRIs. For this application, we use the CoRR
dataset. The application extracts the brain volume of each participant
then computes the average for each group of participants. Unlike the other
applications, the BIDS App Example is a command-line executed in a Docker image
(bids/example on DockerHub). We converted the Docker image to a Singularity
image for use in HPC environments, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}

\subsection{Experiments}
Table~\ref{table:parameters} shows the parameters that we varied
throughout the experiments. We varied (1) the number of nodes to assess
the scalability of the engines' scheduler, (2) the number of files for the BigBrain
in Increment and Histogram to measure the
effect of different I/O patterns and parallelization degrees, and (3) the
number of iterations to evaluate the effect of the number of tasks.
It should be noted that increasing the number of iterations for a given sleep 
delay also increases the total compute time.

@@

\section{Results}
\subsection{Increment}
Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
The bars are broken down into Overhead, Read, Write, and other application-specific functions.
The mean of each bar is calculated from ten repetitions, with the error bar representing the standard deviation.
As expected, the compute time, from \textit{Increment}, remains the same when we vary the number of nodes.
However, the \textit{Read} and \textit{Write} time increases with the number of nodes for both engines.
This is explained by an I/O bottleneck on the Lustre file system.

From Figure \ref{fig:increment_worker}, we also observe that Dask's overhead is higher than Spark's.
While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
Although Dask has a higher overhead, the total time for both engines is overall similar.
This is due to a tradeoff between disk bandwidth and compute/overhead time.
On the one hand, multiple threads access Lustre concurrently, thus sharing the disk bandwidth resulting in longer I/O time.
On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
The only significant difference in total time is when two nodes are used.
This results in an approximately 3\% total runtime difference.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_worker.jpg}
	\caption{Increment: varying nodes -- 5000 files, 5 iterations}
	\label{fig:increment_worker}
\end{figure}

@@

\subsection{Histogram}
Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
Overall, the execution time for Spark is significantly lower than for Dask by up to 5\%, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs.
As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_worker.jpg}
	\caption{Histogram: varying nodes -- 5000 files}
	\label{fig:histogram_worker}
\end{figure}

@@

In a follow-up experiment, we further optimized our Kmeans implementations for both Dask and Spark to minimize the size of data kept in memory.
Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the Gantt chart for the execution
of this new experiments for Dask and Spark.
The x-axis shows the time in second while the y-axis depicts the workers, with 
one worker thread ID (Dask) and process ID (Spark) per line.
We represent each task on the Gantt chart with its associated function's color.
We observe that Spark reports three times more processes than Dask threads.
However, we found out that both Dask and Spark use the same amount of maximum concurrent processes or threads.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=\columnwidth]{figures/dask_kmeans_gantt.png}
	\caption{Kmeans: Dask with 8 nodes, 5000 files, and 3 iterations}
	\label{fig:dask_kmeans_gantt}
\end{figure}

\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=\columnwidth]{figures/spark_kmeans_gantt.png}
	\caption{Kmeans: Spark with 8 nodes, 5000 files, and 3 iterations}
	\label{fig:spark_kmeans_gantt}
\end{figure}

@@

\subsection{BIDS App Example}
Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App Example} application when varying the number of nodes.
We observe that Dask has a slightly lower overhead, thank Spark, although not significant.
However, the makespan for Dask is between 7.5\% to 14.5\% lower than for Spark.
The main difference comes from an increased amount of stagger tasks for Spark, thus increasing the overhead of the application.
Moreover, although minimal, a shorter cluster deployment time for Dask reinforces the difference in overhead time.
As with other experiments, the I/O from Lustre affects the scaling of both engines as we increase the number of nodes.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_bids.jpg}
	\caption{BIDS App Example: varying nodes}
	\label{fig:bids}
\end{figure}

@@

\subsection{Tradeoff between overhead and data transfer}
\label{subsec:overhead-tradeoff}
Overall, except for Kmeans, the overhead for Dask is larger than for Spark.
However, the total execution time for both engines is comparable.
This almost exact compensation between the engine overhead and the data transfer time (see for example Figure \ref{fig:increment_worker}) is not surprising.
When there is high contention on the bandwidth of the file system, and in the absence of trashing, desynchronizing the file transfers generally does not reduce the total data transfer time.
To take an extreme example, the makespan of $n$ concurrent file transfer of size $F$ from a file system with a bandwidth $B$ would be $nF/B$,
as the $n$ transfer would share the bandwidth $B$, which is equivalent to sequentially transferring the $n$ files.
i.e., when the bandwidth of the file system is saturated, engine overhead does not increase the makespan as long as it desynchronizes the data transfers.
In a previous study~\cite{8943502}, we also observed this tradeoff between overhead and data transfer when the file system was saturated.
Exploring strategies that take advantage of this tradeoff could be interesting to reduce the application makespan.

\subsection{Memory management}
Our Kmeans resutls from Figure~{\ref{fig:kmeans_worker}} show a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark using processes compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

@@

\subsection{Scheduling overhead}
Overall, our results show that Dask has a slightly longer startup overhead than Spark.
However, we found that this did not lead directly to the makespan difference.
While both engines showed similar overhead for applications with low inter-worker communication,
Spark had significantly more scheduling overhead than Dask for the Kmeans applications.
Spark was considerably slower to schedule the first set of computations after reading the data on the workers.
Moreover, Spark read all the data before starting computation while was still reading data alongside doing computation.
This difference in scheduling is beneficial for data-intensive applications as it reduces the risk
of I/O bottleneck, as explained in Section~{\ref{subsec:overhead-tradeoff}}.

\subsection{Comparison to other studies}
Overall, except for Kmeans, our experiments did not show substantial performance differences between Spark and Dask.
In contradiction with the experiments of~\cite{Mehta:17}, we found that Dask had a larger overhead than Spark, especially when scaling the number of workers.
However, this was compensated by the previously mentioned overhead and data transfer tradeoff.
Since our focus is on data-intensive applications, we conclude that the performance of Dask and Spark is equivalent.

The experiments from~\cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study~\cite{8588652} found the opposite.
In this study, we found that Dask has a larger startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.

@@

>>>>

\documentclass[AMA,STIX1COL]{WileyNJD-v2}
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage[binary-units=true]{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabularx}

@@

\title{Performance comparison of Dask and Apache Spark on HPC systems \HL{for Neuroimaging}}

@@

\abstract[Summary]{
	The general increase in data size and data sharing motivates the adoption of Big Data strategies in several scientific disciplines.
	However, while several options are available, no particular guidelines exist for selecting a Big Data engine.
	In this paper, we compare the runtime performance of two popular Big Data engines with
	Python APIs, Apache Spark, and Dask, in processing neuroimaging pipelines.
	Our experiments use three synthetic \HL{neuroimaging} applications to process
	the \SI{606}{\gibi\byte} BigBrain image and an actual pipeline to process data
	from thousands of anatomical images.
	We benchmark these applications on a dedicated HPC cluster running the Lustre file system while using varying
	combinations of the number of nodes, file size, and task duration.
	Our results show that although there are slight differences between Dask and Spark,
	the performance of the engines is comparable for data-intensive applications.
	However, Spark requires more memory than Dask, which can lead to 
	slower runtime depending on configuration and infrastructure.
	In general, the limiting factor was the data transfer time.
	While both engines are suitable for neuroimaging, more efforts need to be put
	to reduce the data transfer time and the memory footprint of applications.
}

@@

\section{Introduction}
The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO2018400, van2014human, ConpPortal}.
While standard neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipelines,
they lack Big Data strategies (i.e., in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
A previous study\cite{8752675} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
This work studies how performance benchmarks \HL{of neuroimaging applications} generalize across Big Data engines in an HPC environment.
\HL{
	With the rising need for Big Data engines in neuroimaging and none commonly used
	in the field, it is important to evaluate performance of different engines.
}
In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
\HL{
	Our main contribution is to assess whether Spark or Dask has substantial performance benefits in processing neuroimaging pipelines in HPC systems.
}

@@

While laptops or workstations can be sufficient for some applications, 
neuroimaging pipelines often require large\HL{r} infrastructures
\HL{to process datasets in the scale of several hundreds of {\SI{}{\gibi\byte}} to few {\SI{}{\tebi\byte}}.}
This paper focuses on the performance of Big Data engines in HPCs environments \HL{for neuroimaging}.
We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
Lustre is a reference file system for HPC environments, used in over 60 TOP100 supercomputers~{\cite{OpenSFS-lustre}}, offering parallel file accesses.

In a previous study\cite{8943502} we found no significant performance difference between Apache Spark and Dask for data-intensive applications.
However, using the network file system (NFS) led to an I/O bottleneck limiting application performance.
Another study~\cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
This same study found that the startup overhead time for Spark was more prominent than for Dask.
On the contrary, the work in~\cite{Mehta:17} claims that the startup overhead is more significant for Dask than for Spark.
This contradiction suggests that the startup overhead might vary depending on cluster configuration.
The latter study processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
On the contrary, the work in~\cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
However, that study used the Dask Bag API, which has performance limitations compared to other Dask APIs\HL{{~\cite{DaskBagDoc}}}.

@@

The following section introduces the Lustre file system and the Big Data engines.
Next, we present the design of our benchmarking experiments.
We consider two types of neuroimaging pipelines: high-resolution imaging and functional MRI studies for a cohort of subjects.
The pipelines we chose involve different compute and I/O patterns, e.g., map-reduce or map-only, requiring partial data in memory or the whole dataset.
Further sections present our results, discussion, and conclusion.

@@

\HL{
	While our infrastructure is not of the scale of TOP500 HPC systems,
	it is hard to reserved multiple nodes concurrently on those infrastructure.
	Moreover, by using our infrastructure, we can assure that no external processes
	are influencing the benchmark for our experiements.
	Our system uses similar hardware as Compute Canada's B\'eluga cluster
	--- a representative cluster available to researchers in Canada --- and 
	replicates as close as possible its configuration.
	This lets us compare Spark and Dask in an isolated environement with the condition 
	of a realistic  allocation of resources, for neuroimaging applications, in an HPC system.
}

\HL{
	We configure Dask to have eight worker processes per node, each with eight threads.
	This is the default recommended configuration for the Dask SLURM cluster configuration~{\cite{DaskSLURMDoc}}.
	Similarly, we configured Spark to have eight executors per node, each with eight threads.
	When trying different configuration for the workers, we did not observe performance difference
	for our benchmark.
	Therefore we chose the recommended configuration from Dask.
}
Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
For Spark, the JVM heap space uses 10\% of workers' memory, and the location of the log is set on an NFS-mounted directory.
The worker workspace for both engines is located on Lustre due to the limited fast storage space on the worker nodes.
Other Dask configurations are left to default.
A new Dask or Spark cluster was started and torn down for each experiment.
Dask \textit{2021.7.0} and Spark \textit{3.1.2} were used.

@@

\HL{
	We chose these dataset as they represent common dataset size for neuroimaging 
	imaging studies and for their free access.
}

\subsection{Applications}
This section describes the applications used to benchmark Dask and Spark.
We use three simple synthetic applications to have a deep understanding of the 
underlying behavior of the applications with simple I/O patterns, namely
\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
Also, two neuroimaging applications are used to study more realistic 
applications, namely \textit{Kmeans} and \textit{BIDS App Example}{\cite{gorgolewski2017bids}}.
\HL{
	We chose these neuroimaging data processing applications to study the 
	effects from various Big Data characteristics.
	Table~{\ref{table:applications}} depicts the Big Data characteristics used in
	each application.
}

\begin{table*}[t]
	\caption{Big Data characteristics of the applications}
	\centering
	\begin{tabular}{l|c|c|c|c|c|c|c|}
		\cline{2-8}
		\multicolumn{1}{c|}{} &
		\textbf{Map} &
		\textbf{Reduce} &
		\textbf{\begin{tabular}[c]{@{}c@{}}Data\\ shuffling\end{tabular}} &
		\textbf{\begin{tabular}[c]{@{}c@{}}Dataset \\ in-memory\end{tabular}} &
		\textbf{\begin{tabular}[c]{@{}c@{}}Compute\\ intensive\end{tabular}} &
		\textbf{\begin{tabular}[c]{@{}c@{}}Data\\ intensive\end{tabular}} &
		\textbf{Containerized} \\ \hline
		\multicolumn{1}{|l|}{\textbf{Increment}}        & X &   &         & Partial &   & X &   \\ \hline
		\multicolumn{1}{|l|}{\textbf{Histogram}}        & X & X & Minimal & Partial &   & X &   \\ \hline
		\multicolumn{1}{|l|}{\textbf{Kmeans}}           & X &   & X       & Whole   & X &   &   \\ \hline
		\multicolumn{1}{|l|}{\textbf{BIDS App Example}} & X & X & Minimal & Partial & X &   & X \\ \hline
	\end{tabular}
	\label{table:applications}
\end{table*}

\HL{
	To minimize the performance difference from implementation decisions, the 
	code base of the applications is common for both engines.
	Therefore, the performance observed comes directly from the engine scheduling and
	their APIs.
}

@@

\subsubsection{Histogram}
As our second application, we calculate the histogram of the BigBrain image. 
The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre (Fig.~\ref{fig:graph-histogram}).
This map-reduce application has a very high read over write ratio.
Moreover, this application requires shuffling, albeit of a limited amount of data. 
The amount of inter-worker communication is in-between the increment and multi-increment applications.

\begin{figure}[h]
	\centering
	\begin{minipage}{0.4\textwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/increment.png}
		\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
		\label{fig:graph-increment}		
	\end{minipage}
	\begin{minipage}{0.55\textwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/histogram.png}
		\caption{Task graph for Histogram with 3 BigBrain blocks.}
		\label{fig:graph-histogram}			
	\end{minipage}
\end{figure}

\subsubsection{Kmeans}
For our third application, we apply Kmeans clustering to the voxel
intensities of the BigBrain image. We set the number of clusters to 3. The
application starts by reading the image blocks, combining all voxels in a
1-D array, and choosing initial centroids using the min, max, and
intermediate values. It assigns each voxel to its closest centroid and
updates each centroid with the average of its assigned voxels. It repeats
the assignment and updates steps for a configurable number of iterations.
Finally, the voxels of the image blocks are classified and written back to
the file system. Updating the centroids involves substantial data
communication between the workers.
The computation to find the closest centroids for each voxels scales with the number of centroids.
It can be computed in parallel at the cost of higher memory or sequentially to reduce memory usage.
We decided to compute the nearest centroids in parallel since we use only three centroids.
However, due to resources limitation on our infrastructure, the Kmeans experiments
are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).
\HL{We chose this applications for its real-life use case, large data shuffling between workers,
and requirements to keep all data in-memory.}

\subsubsection{BIDS App Example}
Our fourth application is the BIDS App Example, a neuroimaging pipeline to
measure brain volume from MRIs. For this application, we use the CoRR
dataset. The application extracts the brain volume of each participant
then computes the average for each group of participants. Unlike the other
applications, the BIDS App Example is a command-line executed in a Docker image
(bids/example on DockerHub). We converted the Docker image to a Singularity
image for use in HPC environments, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
\HL{We chose this application for its real-life use case and it is containerized unlike our applications.}

\subsection{Experiments}
Table~\ref{table:parameters} shows the parameters that we varied
throughout the experiments. We varied (1) the number of nodes to assess
the scalability of the engines' scheduler, (2) the number of files \HL{to segment}
BigBrain to measure the effect of different I/O patterns and parallelization degrees, and (3) the
number of iterations to evaluate the effect of the number of tasks.
It should be noted that increasing the number of iterations for a given sleep 
delay also increases the total compute time.
\HL{
	The choice of parameters for our experiements was restrained by the datasets
	and resources available from the cluster.
	The unrestrained parameters were chosen to have an expected linear scaling 
	to facilitate the interpretation of the results.
}

@@

\section{Results}
\subsection{Increment}
Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
The bars are broken down into Overhead, Read, Write, and other application-specific functions.
The mean of each bar is calculated from ten repetitions, with the error bar representing the standard deviation.
As expected, the compute time, from \textit{Increment}, remains the same when we vary the number of nodes.
However, the \textit{Read} and \textit{Write} time increases with the number of nodes for both engines.
\HL{
	With a large amount of worker threads accessing the Lustre file system
	concurrently, the disk bandiwdth or network are saturated and it create 
	an I/O bottleneck to access the Lustre file system.
}

From Figure \ref{fig:increment_worker}, we also observe that Dask's overhead is higher than Spark's.
While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
Although Dask has a higher overhead, the total time for both engines is overall similar.
This is due to a tradeoff between disk bandwidth and compute/overhead time.
\HL{
	On the one hand, when multiple threads access Lustre concurrently they risk 
	of saturation the shared disk bandwidth, thus resulting in longer I/O time.
	On the other hand, when most threads are either idle or performing computations,
	it reduces the contention on the disk bandwidth resulting in shorter I/O time.
}
The only significant difference in total time is when two nodes are used.
This results in an approximately 3\% total runtime difference.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_worker.jpg}
	\caption{Increment: varying nodes -- 5000 files, 5 iterations}
	\label{fig:increment_worker}
\end{figure}

@@

\subsection{Histogram}
Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
Overall, the execution time for Spark is significantly lower than for Dask by up to 5\%, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs\HL{~{\cite{DaskBagDoc}}}.
As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_worker.jpg}
	\caption{Histogram: varying nodes -- 5000 files}
	\label{fig:histogram_worker}
\end{figure}

@@

In a follow-up experiment, we further optimized our Kmeans implementations for both Dask and Spark to minimize the size of data kept in memory.
Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the Gantt chart for the execution
of this new experiments for Dask and Spark.
The x-axis shows the time in second while the y-axis depicts the workers, with 
one worker thread ID (Dask) and process ID (Spark) per line.
We represent each task on the Gantt chart with its associated function's color.
We observe that Spark reports three times more processes than Dask threads.
\HL{
	It is unclear what is the cause for reporting extra Spark processes.
	However, with our analysis we found that both Dask and Spark use the same 
	amount of maximum concurrent processes or threads.
	Additionally, the is no significant difference in the average amount of 
	concurrent processes or threads between Dask and Spark.
}

@@

\begin{figure}[!h]
	\centering
	\href{https://mathdugre.me/paper-big-data-engines/ccpe/dask_kmeans_gantt.html}
	{\includegraphics[clip,width=\columnwidth]{figures/dask_kmeans_gantt.png}}
	\caption{Kmeans: Dask with 8 nodes, 5000 files, and 3 iterations}
	\label{fig:dask_kmeans_gantt}
\end{figure}

\begin{figure}[!h]
	\centering
	\href{https://mathdugre.me/paper-big-data-engines/ccpe/spark_kmeans_gantt.html}
	{\includegraphics[clip,width=\columnwidth]{figures/spark_kmeans_gantt.png}}
	\caption{Kmeans: Spark with 8 nodes, 5000 files, and 3 iterations}
	\label{fig:spark_kmeans_gantt}
\end{figure}

@@

\subsection{BIDS App Example}
Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App Example} application when varying the number of nodes.
We observe that Dask has a slightly lower overhead, thank Spark, although not significant.
However, the makespan for Dask is between 7.5\% to 14.5\% lower than for Spark.
The main difference comes from an increased amount of stagger tasks for Spark, thus increasing the overhead of the application.
Moreover, although minimal, a shorter cluster deployment time for Dask reinforces the difference in overhead time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_bids.jpg}
	\caption{BIDS App Example: varying nodes}
	\label{fig:bids}
\end{figure}

@@

\subsection{Tradeoff between overhead and data transfer}
\label{subsec:overhead-tradeoff}
Overall, except for Kmeans, the overhead for Dask is larger than for Spark.
However, the total execution time for both engines is comparable.
This almost exact compensation between the engine overhead and the data transfer time (see for example Figure \ref{fig:increment_worker}) is not surprising.
When there is high contention on the bandwidth of the file system, and in the absence of trashing, desynchronizing the file transfers generally does not reduce the total data transfer time.
To take an extreme example, the makespan of $n$ concurrent file transfer of size $F$ from a file system with a bandwidth $B$ would be $nF/B$,
as the $n$ transfer would share the bandwidth $B$, which is equivalent to sequentially transferring the $n$ files.
In a previous study~\cite{8943502}, we also observed this tradeoff between overhead and data transfer when the file system was saturated.
Exploring strategies that take advantage of this tradeoff could be interesting to reduce the application makespan.

\subsection{Memory management}
Our Kmeans results from Figure~{\ref{fig:kmeans_worker}} show a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark using processes compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

@@

\subsection{Scheduling overhead}
Overall, our results show that Dask has a slightly longer startup overhead than Spark.
However, we found that this did not lead directly to the makespan difference.
While both engines showed similar overhead for applications with low inter-worker communication,
Spark had significantly more scheduling overhead than Dask for the Kmeans applications.
Spark was considerably slower to schedule the first set of computations after reading the data on the workers.
Moreover, Spark read all the data before starting computation while Dask was still reading data alongside doing computation.
This difference in scheduling is beneficial for data-intensive applications as it reduces the risk
of I/O bottleneck, as explained in Section~{\ref{subsec:overhead-tradeoff}}.

\subsection{Comparison to other studies}
Overall, except for Kmeans, our experiments did not show substantial performance differences between Spark and Dask.
In contradiction with the experiments of~\cite{Mehta:17}, we found that Dask had a larger overhead than Spark, especially when scaling the number of workers.
However, this was compensated by the previously mentioned overhead and data transfer tradeoff.
Since our focus is on \HL{neuroimaging} applications, we conclude that the performance of Dask and Spark is equivalent.

\HL{
	In this study, we observed that Dask has a larger startup overhead that Spark.
	These results are consistent with a previous study~{\cite{Mehta:17}}.
	However, other studies found the opposite~{\cite{8588652}} or that the difference was negligible~{\cite{8943502}}
}
Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.

@@

\subsection{Recommendations}
\iffalse
The differences in performance are minimal between both engines. However, 
we added a “Recommendation” subsection to the discussion for qualitative 
difference and some caveats to avoid when using the engines.
\fi
\HL{
	The results from our experiments show that the performance difference between
	Dask and Spark is minimal, in general. 
	The main differences we observed are qualitative.
	We found that Dask integration with the scientific Python ecosystem made it
	easier to implement neuroimaging application.
	Spark configuration required to set multiple environment variables.
	In comparison, Dask is ready-to-go after an installation through PIP.
	Dask dashboard provides more information than Spark ones, making
	the debugging of Dask application easier and faster.
	Therefore, for the neuroimaging field, we recommend the use of Dask over Spark.
	However, in other fields with non-Python-based ecosystem, both Dask and spark
	would be equal candidate as a Big Data engine.
}

\HL{
	% TODO add caveats to avoid when using the engines.
}
